Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-11 paper.tex
Abbreviate figure references.

The constructionist learning philosophy emphasizes creation and
tangible interaction.  Nell uses direct interaction on a tablet
computer to maintain the child's connection to their work
(Figure~\ref{fig:xo3}).

@@

\subsection{Rule-based story modules}
\begin{figure}\small
\begin{verbatim}
{ world: { // id gives canonical URL for this module
    id: "/nell.laptop.org/chapters/alphabet-book",
    scene: "opening", // initial scene
    inherit_from: "/nell.laptop.org/core/drawing_activity",
  }, scenes: [
    { id: "opening", subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Hello! Would (o/2) like to hear a story?",
      group: "/nell.laptop.org/core/story-intro-group",
    },
    { id: "story-1",
      cond: rule`At(opening) && Event(Chat, Yes)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Once upon a time..." },
    ...
    { id: "r-drawn",
      cond: rule`Event(Activity, DrawR)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Very good, (o/3)!",
      action: js`startAnimation("r-runs")` }, ...
  ], ... }
\end{verbatim}
\caption{Excerpt of story module.  JSON syntax has been extended with
quasiquote~\cite{quasiquote}.}\label{fig:rules}
\end{figure}
To support Nell's \textit{modular narrative interface}, we use a rule-based
story model with programmable conflict resolution. (Figure~\ref{fig:rules} shows
an excerpt of the story module underlying the interaction described in
Section~\ref{sec:lesson}.)

@@

Conflicts are resolved by consulting a \textit{group} associated with
each scene.  Each group has a default priority and names a scene to be
invoked when a conflict exists; the actions associated with that scene
will temporarily alter the priorities of the conflicting scenes to
resolve the conflict.  For example, the \textit{story-intro-group}
cited in the example defaults to a lower priority, such that
continuing a story in progress is preferred to beginning a new scene.
The resolution scene for \textit{story-intro-group} may be part of its
own story: for example, it may bring us to the controls of a magic airship
which allows us to pilot between locations corresponding to the
various story continuations.  Selecting a destination location bumps
the priority of the given scene allowing us to transition to it
without further conflict.

@@

Figure~\ref{fig:aiml}
shows how this can be used to implement a traditional
text adventure interaction for a particular story module.
In this example the child can write instructions such as, ``go north,'' ``go
east,'' or ``open door'' for the agent.  These trigger events
handled by the story module to move the agent through the
adventure---for example, by causing the activity to draw a
new location, causing the agent to describe it, and shifting the
AIML topic to enable vocabulary suitable for the new location.

>>>>

The constructionist learning philosophy emphasizes creation and
tangible interaction.  Nell uses direct interaction on a tablet
computer to maintain the child's connection to their work
(Fig.~\ref{fig:xo3}).

@@

\subsection{Rule-based story modules}
\begin{figure}\small
\begin{verbatim}
{ world: { // id gives canonical URL for this module
    id: "/nell.laptop.org/chapters/alphabet-book",
    scene: "opening", // initial scene
    inherit_from: "/nell.laptop.org/core/drawing_activity",
  }, scenes: [
    { id: "opening", subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Hello! Would (o/2) like to hear a story?",
      group: "/nell.laptop.org/core/story-intro-group",
    },
    { id: "story-1",
      cond: rule`At(opening) && Event(Chat, Yes)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Once upon a time..." },
    ...
    { id: "r-drawn",
      cond: rule`Event(Activity, DrawR)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Very good, (o/3)!",
      action: js`startAnimation("r-runs")` }, ...
  ], ... }
\end{verbatim}
\caption{Excerpt of story module.  JSON syntax has been extended with
quasiquote~\cite{quasiquote}.}\label{fig:rules}
\end{figure}
To support Nell's \textit{modular narrative interface}, we use a rule-based
story model with programmable conflict resolution. (Fig.~\ref{fig:rules} shows
an excerpt of the story module underlying the interaction described in
Section~\ref{sec:lesson}.)

@@

Conflicts are resolved by consulting a \textit{group} associated with
each scene.  Each group has a default priority and names a scene to be
invoked when a conflict exists; the actions associated with that scene
will temporarily alter the priorities of the conflicting scenes to
resolve the conflict.  For example, the \textit{story-intro-group}
used in Fig.~\ref{fig:rules} defaults to a lower priority, such that
continuing a story in progress is preferred to beginning a new scene.
The resolution scene for \textit{story-intro-group} may be part of its
own story: for example, it may bring us to the controls of a magic airship
which allows us to pilot between locations corresponding to the
various story continuations.  Selecting a destination location bumps
the priority of the given scene allowing us to transition to it
without further conflict.

@@

Fig.~\ref{fig:aiml}
shows how this can be used to implement a traditional
text adventure interaction for a particular story module.
In this example the child can write instructions such as, ``go north,'' ``go
east,'' or ``open door'' for the agent.  These trigger events
handled by the story module to move the agent through the
adventure---for example, by causing the activity to draw a
new location, causing the agent to describe it, and shifting the
AIML topic to enable vocabulary suitable for the new location.
